A Theory for Data Migration of Information Systems

### 2022-09-17 2022 Migration/articleMigration.tex
Text erbij over de ENFORCE en CLASSIFY regels

	We use two datasets: $\dataset=\la\atoms,\concepts,\inst,\isa,\rels,\triples\ra$ and $\dataset'=\la\atoms',\concepts',\inst',\isa',\rels',\triples'\ra$.
	Before doing so, let us first define the disjoint union of two information systems.
\begin{definition}[disjoint union of datasets]
	\begin{eqnarray}
		\dataset\sqcup\dataset'&=&\la\atoms\uplus\atoms',\ \concepts\uplus\concepts',\ \inst\uplus^2\inst',\ \isa\uplus^2\isa',\ \rels\uplus^3\rels',\ \triples\uplus^5\triples'\ra\\
		X\uplus Y&=&\{(x,0)|\ x\in X\}\ \cup\ \{(y,1)|\ y\in Y\}\\
		X\uplus^2 Y&=&\begin{array}[t]{@{}l}\{((x_1,0),(x_2,0))|\ (x_1,x_2)\in X\}\ \cup\\ \{((y_1,1),(y_2,1))|\ (y_1,y_2)\in Y\}\end{array}\\
		X\uplus^3 Y&=&\begin{array}[t]{@{}l}\{\declare{(n,0)}{(A,0)}{(B,0)}\mid\ \declare{n}{A}{B}\in X\}\ \cup\\ \{\declare{(n,1)}{(A,1)}{(B,1)}\mid\ \declare{n}{A}{B}\in Y\}\end{array}\\
		X\uplus^5 Y&=&\begin{array}[t]{@{}l}\{\triple{(a,0)}{\declare{(n,0)}{(A,0)}{(B,0)}}{(b,0)}\mid\ \declare{n}{A}{B}\in X\}\ \cup\\ \{\triple{(a,1)}{\declare{(n,1)}{(A,1)}{(B,1)}}{(b,1)}\mid\ \declare{n}{A}{B}\in Y\}\end{array}
	\end{eqnarray}
\end{definition}

\begin{definition}[migration dataset]
	\begin{eqnarray}
		&&\begin{array}[t]{@{}l}
			\{ {\tt ENFORCE\ }\declare{(nm',1)}{(A',1)}{(B',1)}\\\quad{\tt := }\ident{(A',1)};\declare{(nm,0)}{(A,0)}{(B,0)};\ident{(B',1)}\\
				\mid\ \begin{array}[t]{@{}l}
					\declare{nm}{A}{B}\in\rels\wedge\declare{nm'}{A'}{B'}\in\rels'\wedge\id{nm}=\id{nm'}\ \wedge\\
					\{\pair{A}{A'},\pair{B}{B'}\}\subseteq\isa\cup\isa'\cup\flip{\isa}\cup\flip{\isa'}\}
				   \end{array}
			  \end{array}\\
			  &\cup&\begin{array}[t]{@{}l}
				\{ {\tt CLASSIFY\ }(C',1){\tt\ IS\ }(C,0)|\ C\in\concepts,C'\in\concepts',C=C'\}\\
			  \end{array}
	\end{eqnarray}
\end{definition}
\begin{definition}[disjoint union of information systems]
	\begin{eqnarray}
		\infsys\sqcup\infsys'&=&\la\roles\uplus\roles',\rules\uplus^?\rules',\maintain\uplus^2\maintain',\dataset\sqcup\dataset'\ra
	\end{eqnarray}
\end{definition}

>>>>

	We use two datasets: $\dataset=\la\atoms,\concepts,\inst,\isa,\rels,\triples\ra$ and $\dataset'=\la\atoms',\concepts',\inst',\isa',\rels',\triples'\ra$.
	Before doing so, let us first define the disjoint union of two information systems.
\begin{definition}[disjoint union of datasets]
	\begin{eqnarray}
		\dataset\sqcup\dataset'&=&\la\atoms\uplus\atoms',\ \concepts\uplus\concepts',\ \inst\uplus^2\inst',\ \isa\uplus^2\isa',\ \rels\uplus^3\rels',\ \triples\uplus^5\triples'\ra\\
		X\uplus Y&=&\{(x,0)|\ x\in X\}\ \cup\ \{(y,1)|\ y\in Y\}\\
		X\uplus^2 Y&=&\begin{array}[t]{@{}l}\{((x_1,0),(x_2,0))|\ (x_1,x_2)\in X\}\ \cup\\ \{((y_1,1),(y_2,1))|\ (y_1,y_2)\in Y\}\end{array}\\
		X\uplus^3 Y&=&\begin{array}[t]{@{}l}\{\declare{(n,0)}{(A,0)}{(B,0)}\mid\ \declare{n}{A}{B}\in X\}\ \cup\\ \{\declare{(n,1)}{(A,1)}{(B,1)}\mid\ \declare{n}{A}{B}\in Y\}\end{array}\\
		X\uplus^5 Y&=&\begin{array}[t]{@{}l}\{\triple{(a,0)}{\declare{(n,0)}{(A,0)}{(B,0)}}{(b,0)}\mid\ \declare{n}{A}{B}\in X\}\ \cup\\ \{\triple{(a,1)}{\declare{(n,1)}{(A,1)}{(B,1)}}{(b,1)}\mid\ \declare{n}{A}{B}\in Y\}\end{array}
	\end{eqnarray}
\end{definition}
Note that if $\dataset$ and $\dataset'$ are datasets, then so is $\dataset\sqcup\dataset'$.

\begin{definition}[migration dataset]
	\begin{eqnarray}
		&&\begin{array}[t]{@{}l}
			\{ {\tt ENFORCE\ }\declare{(nm',1)}{(A',1)}{(B',1)}\\\quad{\tt\ :=\ }\ident{(A',1)};\declare{(nm,0)}{(A,0)}{(B,0)};\ident{(B',1)}\\
				\mid\ \begin{array}[t]{@{}l}
					\declare{nm}{A}{B}\in\rels\wedge\declare{nm'}{A'}{B'}\in\rels'\wedge\id{nm}=\id{nm'}\ \wedge\\
					\{\pair{A}{A'},\pair{B}{B'}\}\subseteq\isa\cup\isa'\cup\flip{\isa}\cup\flip{\isa'}\}
				   \end{array}
			  \end{array}\\
			  &\cup&\begin{array}[t]{@{}l}
				\{ {\tt CLASSIFY\ }(C',1){\tt\ IS\ }(C,0)|\ C\in\concepts,C'\in\concepts',C=C'\}\\
			  \end{array}
	\end{eqnarray}
\end{definition}

The ${\tt CLASSIFY\ }(C',1){\tt\ IS\ }(C,0)$ statements add $((C',1),(C,0))$ and  $((C,0),(C',1))$ to the $\isa$ relation. Note that if $\dataset$ is a dataset, and $\dataset'$ is that dataset with pairs added\footnote{TODO: het woord `add' is informeel, dit kan beter} to the $\isa$ relation, then the result is again a dataset.
The ${\tt ENFORCE\ }r{\tt\ :=\ }e$ statements add triples $(x,r,y)$ for all $(x,y)\in e$.
The expressions $e$ in these statements are such that $x\ (\inst \compose \kleenestar{\flip{\isa}})\ (A',1)$ and $y \ (\inst \compose \kleenestar{\flip{\isa}})\ (B',1)$, thus preserving the property that the result is a dataset.

\begin{definition}[disjoint union of information systems]
	\begin{eqnarray}
		\infsys\sqcup\infsys'&=&\la\roles\uplus\roles',\rules\uplus^?\rules',\maintain\uplus^2\maintain',\dataset\sqcup\dataset'\ra
	\end{eqnarray}
\end{definition}
